Measuring Perceptions and Preferences for Meritocracy

### 2023-02-02 paper-blinded.Rmd
grammarly vscode check

---
title: "Measur---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items. Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies. The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items. Based on these results, in a second stage we propose  an  improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141). The results of the CFA estimation give evidence of a better fit for the new scale. Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

@@

Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and (or) contextual characteristics. For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019]. Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy. However, as we will show later, the concepts and instruments used for measuring meritocracy vary extensively among studies. In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators. Such trends limit the comparability of studies and the ability to understand the effects of meritocratic beliefs across different disciplines and lines of research.

Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences. Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies suggest. This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.

@@

In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy? Whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence. Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020]. Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations. To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above. Some versions of this battery ask "how important you think it is for getting ahead in life" and then lists several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.  Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992. As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, beliefs use to cover almost anything related to subjective factors. To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (i.e. what should be), while referring to perceptions (i.e. what is). For instance,  in @reynolds_perceptions_2014, the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007]. The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs). They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "... people are rewarded for their skills?".

@@

Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life. This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.  For instance, @kunovich_systems_2007 used an items' battery following the question “How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy). They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued as 1 when rated as "not important at all" (i.e. reverse coded). The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy. A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic. Despite making this distinction, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did. Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment. For instance, such distinction could open the door to study to what extent different and apparently contradictory views of meritocracy can coexist, to what extent individuals and societies differ to this regard, and of course the associated factors to such differences. Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.

@@

There is a relevant trade-off regarding measurement error estimation in survey scales. Including multiple items per conceptual dimension opens the possibility to validity studies (at least in terms of construct) by using factor analysis techniques, particularly confirmatory ones. However, increasing the amount of items reduces the probability of usability of the scale in survey research given space limitations. This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted by a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.

@@

The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019]. Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society". Preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research. Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to getting ahead in life.

@@

The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic). There were two items from the battery that were excluded from the analysis as they would not fit into the classification. The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.

@@

In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix. Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp). The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections). Still, between this same set of items, moderate correlations (around 0.3) could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.

@@

Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models. The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535). Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.

@@

Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions. Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum. Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).

@@

## Discussion study 1

Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy. First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but it still is possible to find a certain fit of the model to the data after minor adjustments. Secondly, some of the items in similar dimensions refer to different concepts, as for instance meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done. A better operationalization would require more alignment between the different sides of the same concept. And thirdly, the scale is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors. All these limitations are taken into account next in the design of Study 2.

@@

The data comes from an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID). The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta). The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not sign the informed consent and those not answering the meritocracy response battery. There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \@ref(tab:rep-samp) in Appendix). As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020]. This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.

@@

The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). Two main criteria oriented this design. The first one was usability in social survey research, for which the scale length must attend spacer restrictions in such studies. Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques. Although a rule of thumb in latent measurement suggests at least three indicators per dimension, we still opted for two to count with a brief scale. Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).

The second criteria is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area. To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014]. The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores). These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).

@@

With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod). The scale was presented to the first group (_n = 712_) in the order shown in Table \@ref(tab:table-indicadores) according to perceptions and preferences. For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics. Finally, for the third group (_n = 712_), the items were presented as completely randomized.

@@

## Discussion study 2

The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model. The four dimensions appear independent but related, with some differences when compared with the estimation with ISSP data in Study 1. The main general difference is that with ISSP data, the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero. For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair. Theoretically, this makes more sense, as if someone perceives that meritocracy functions well, they probably will support it more. On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to work.

@@

The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model. Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012]. Besides, the literature on measurement invariance suggests some complementary approaches for evaluamerit scale the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].

@@

Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation. When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal. Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012]. Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$. Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007]. Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.

@@

Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples. This gives a more robust ground when using this measure for comparing groups, as for instance in country comparisons. Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items. Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual and measurement framework for perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies. The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items. Based on these results, in a second study we propose an  improved scale that better suits the perceptions/preferences distinction and that is tested  in a Chilean sample (N=2,141). The results of the CFA estimation give evidence of a better fit for the new scale. Our discussion highlights the importance of considering different dimensions in the study of meritocracy, both conceptually and empirically."
output:
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

@@

Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and (or) contextual characteristics. For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019]. Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy. However, as we will show later, the concepts and instruments used for measuring meritocracy vary extensively among studies. In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators. Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research.

Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences. Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest. This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.

@@

In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy? We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence in studies focused on the subjective aspects of inequality. Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, for instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020]. Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort as in the original Young's conception.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations. To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above. Some versions of this battery ask "how important you think it is for getting ahead in life" and then list several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.  Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992. As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, beliefs use to cover almost anything related to subjective factors. In this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e., thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (i.e. what should be), while referring to perceptions (i.e., what is). For instance,  in @reynolds_perceptions_2014, the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007]. The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs). They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "... people are rewarded for their skills?".

@@

Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life. This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.  For instance, @kunovich_systems_2007 used an items' battery following the question “How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy). They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued as 1 when rated as "not important at all" (i.e. reverse coded). The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (such as having family and/or children) implies a stronger belief in meritocracy. A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic. Despite making this distinction, in the operationalization the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did. Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment. For instance, a such distinction could open the door to studying to what extent different and contradictory views of meritocracy can coexist, to what extent individuals and societies differ in this regard, and of course the associated factors to such differences. Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of other terms like ascription, given that this last one leaves out aspects such as the use of personal relationships as a way to achieve success.

@@

There is a relevant trade-off regarding measurement error estimation in survey scales. Including multiple items per conceptual dimension opens the possibility of validity studies (at least in terms of the construct) by using factor analysis techniques, particularly confirmatory ones. However, increasing the number of items reduces the probability of usability of the scale in survey research given space limitations. This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.

@@

The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019]. Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society". Preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research. Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to get ahead in life.

@@

The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic). There were two items from the battery that were excluded from the analysis as they would not fit into the classification. The first one was "having good education yourself" since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.

@@

In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix. Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp). The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections). Still, between this same set of items, moderate correlations (around 0.3) could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.

@@

Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models. The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535). Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.

@@

Regarding the correlations between the factors in Figure ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions. Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum. Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).

@@

## Discussion Study 1

Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy. First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but it still is possible to find a certain fit of the model to the data after minor adjustments. Secondly, some of the items in similar dimensions refer to different concepts, for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done. A better operationalization would require more alignment between the different sides of the same concept. And thirdly, the scale is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors. All these limitations are taken into account next in the design of Study 2.

@@

The data comes from an online survey that was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID). The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta). The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not sign the informed consent and those not answering the meritocracy response battery. There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \@ref(tab:rep-samp) in Appendix). As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020]. This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.

@@

The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). Two main criteria oriented this design. The first one was usability in social survey research, for which the scale length must attend spacer restrictions in such studies. Nevertheless, at the same time, the scale should offer possibilities of accounting for measurement error through factor analysis techniques. Although a rule of thumb in latent measurement suggests at least three indicators per dimension, we still opted for two to count with a brief scale. Although such a decision posits some limitations for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009). The second criterion is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area. To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014]. The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores). These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).

@@

To evaluate the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod). The scale was presented to the first group (_n = 712_) in the order shown in Table \@ref(tab:table-indicadores) according to perceptions and preferences. For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics. Finally, for the third group (_n = 712_), the items were presented as completely randomized.

@@

## Discussion Study 2

The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model. The four dimensions appear independent but related, with some differences when compared to the estimation with ISSP data in Study 1. The main general difference is that with ISSP data, the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero. For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair. Theoretically, this makes more sense, as if someone perceives that meritocracy functions well, they probably will support it more. On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to work.

@@

The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model. Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012]. Besides, the literature on measurement invariance suggests some complementary approaches for evaluating the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].

@@

Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation. When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal. Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012]. Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$. Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007]. Therefore, using this last standard, the level of strict invariance would hold for the meritocracy scale.

@@

Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples. This gives a more robust ground when using this measure for comparing groups, for instance in country comparisons. Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.
